---
title: "Integrated_framework_scRNA-seq_annotation"
author: "Saman Farahmand"
date: "01/10/2022"
output: html_document
---

## R Markdown
# Analysis tracking

+ Author: [Saman Farahmand](mailto:saman.farahmand@takeda.com)
+ Start Analysis: 12012021
+ Update Analysis: 01102022
+ Status: Complete 


```{r echo=FALSE, warning=FALSE}
library(slingshot)
library(BUSpaRse)
library(tidyverse)
library(tidymodels)
library(Seurat)
library(scales)
library(viridis)
library(Matrix)
library(ggplot2)
library(ggthemes)
library(ggbeeswarm)
library(scmap)
library(SingleR)
library(dplyr)
library(scclusteval)
library(SCINA)
library(data.table)
library(scran)
library(caret)

# loading utility functions
source("functions_util.R")

```


###  Loading Smillie Training and Testing sets: Fibroblast/Epithelium/Immune

```{r}

# loading the training set
train_objects = readRDS("files/Smillie_Stromal_folds/train_objects.RDS")
test_objects = readRDS("files/Smillie_Stromal_folds/test_objects.RDS")

```



## Set the flags according to the annotation methods you want to run:

```{r}
SCINA = T
scmap_cell = T
scmap_cluster = T
```

# Implementing the annotation workflow:

We need to construct a script to get train and test seurat objects and return the annotated test object as a seurat object:

There are multiple methods implemented and we can decide what mthods we want to run here:

1 - SCINA = True/False # If this parameter is True, we need to feed a list of markers named with the labels

2 - scmap

4 - SingleR: # with or without markers


### Running SCINA method


### Preparing the markers for SCINA: 

1 - For Epithelium cells:
```{r}

# Import original markers reported by Smillie et. al.
markers_smillie_all = read.delim("files/smillie_markers_orig.txt",header = T, sep = "\t")
colnames(markers_smillie_all)
label_all=colnames(markers_smillie_all)
markers = as.list(label_all)
names(markers) = label_all
for (marker in markers){
    markers[[marker]] = markers_smillie_all[[marker]][markers_smillie_all[[marker]]!=""]
}
# names is optional: for syncing the labels with other references we have
# It should be consistent with the order of columns in the marker table

## For Epithelial cells
names = c("Enterocyte Progenitors", "Immature Enterocytes 1",  "Immature Enterocytes 2",  "Enterocytes","Best4+ Enterocytes","Stem","Secretory TA", "Cycling TA","TA 1","TA 2","Enteroendocrine", "Goblet", "Immature Goblet","M cells","Tuft")


names(markers) = names


# Import refined set of markers
markers_smillie_refined = read.csv("files/geneSets_flt_markers_refined_CD8_IELs.txt",header = T, sep = "\t")
markers_smillie_refined = markers_smillie_refined %>% dplyr::select(c("ident", "gene"))
smillie_celltypes=unique(markers_smillie_refined$ident)
smillie_markers = as.list(smillie_celltypes)
names(smillie_markers) = smillie_celltypes

for (cell in smillie_markers){
  smillie_markers[[cell]] = markers_smillie_refined[markers_smillie_refined$ident==cell,"gene"]
}

cell_types = c("Goblet", "Immature Goblet", "Goblet", "Tuft", "TA", "Secretory TA", "Cycling TA", "Enteroendocrine", "Stem", "TA 2", "TA 1", "Enterocytes","Enterocyte Progenitors", "M cells", "Best4+ Enterocytes", "Immature Enterocytes 1" , "Immature Enterocytes 2")
## Filter in only Epithelial cells
smillie_markers = smillie_markers[names(smillie_markers) %in% cell_types ]



```

2 - For Fibroblast cells:

```{r}

# Import original markers reported by Smillie et. al.
markers_smillie_all = read.delim("files/Fibro_markers.csv",header = T, sep = ",")
colnames(markers_smillie_all)
label_all=colnames(markers_smillie_all)
markers = as.list(label_all)
names(markers) = label_all
for (marker in markers){
    markers[[marker]] = markers_smillie_all[[marker]][markers_smillie_all[[marker]]!=""]
}
# names is optional: for syncing the labels with other references we have
# It should be consistent with the order of columns in the marker table

## For Epithelial cells
names = c("Myofibroblasts", "WNT2B+ Fos-hi",  "WNT2B+ Fos-lo 1",  "WNT2B+ Fos-lo 2","RSPO3+","WNT5B+ 1","WNT5B+ 2", "Inflammatory Fibroblasts","Endothelial","Microvascular","Post-capillary Venules", "Pericytes")

names(markers) = names



```

3 - For Immune cells:
```{r}
# Import original markers reported by Smillie et. al.
markers_smillie_all = read.delim("files/immune_markers.txt",header = T, sep = "\t")
colnames(markers_smillie_all)
label_all=colnames(markers_smillie_all)
markers = as.list(label_all)
names(markers) = label_all
for (marker in markers){
    markers[[marker]] = markers_smillie_all[[marker]][markers_smillie_all[[marker]]!=""]
}
# names is optional: for syncing the labels with other references we have
# It should be consistent with the order of columns in the marker table

## For Epithelial cells
names = c("DC1",	"DC2",	"CD69+ Mast",	"CD69- Mast",	"Macrophages",	"Cycling Monocytes",	"Inflammatory Monocytes",	"NKs",	"ILCs",	"Cycling T",	"Tregs", "CD4+ Activated Fos-hi",	"CD4+ Activated Fos-lo",	"CD4+ Memory",	"CD4+ PD1+",	"MT-hi",	"CD8+ IELs",	"CD8+ IL17+",	"CD8+ LP",	"Follicular",	"GC",	"Cycling B")

names(markers) = names



```



```{r}


# with original markers
if(SCINA) {
  for(i in c(1:length(test_objects))){
      test_objects[[i]] = Run_SCINA(test_objects[[i]],markers = markers)
      test_objects[[i]]$SCINA_orig = test_objects[[i]]$SCINA
      
  }
}

```


## Run scmap cluster and cell methods


```{r}

for(i in c(1:length(test_objects))){
      test_objects[[i]] = Run_Scmap(train_objects[[i]],test_objects[[i]],ref_assay ="RNA", query_assay = "RNA",
                                    scmap_cell = T,
                                    scmap_cluster = T,
                                    n_features=500, Cluster="Cluster")
}

```



# Save the test and train objects 
 We'll use test_object for SingleR runs separately

```{r}
saveRDS(test_objects,"files/Smillie_Stromal_folds/test_objects.RDS")
saveRDS(train_objects,"files/Smillie_Stromal_folds/train_objects.RDS")


```



## Running SingleR algorithm. It has three versions:
1 - without markers
2 - with list of markers corresponding to each cell type
3 - with DE markers corresponding to each paiwise comparisons

# You can run the script by selecting one of these options or all together

```{r}

######## You need to run SingleR_CV.R script 

```

## Integrating all the annotations in a single object:

```{r}

seur_SRde=readRDS("files/Smillie_Stromal_folds/test_objects_de.RDS")

for (i in c(1:length(test_objects))){
  test_objects[[i]]$singleRde = seur_SRde[[i]]$singleRDE
}

```


# let's imlement the majority voting

```{r echo=FALSE, warning=FALSE}

annotation_columns = c("SCINA_orig","singleRde","scPred","Azimuth")

for (i in c(1:5)){
  test_objects[[i]]$consensus_lab = apply(test_objects[[i]]@meta.data[,annotation_columns], 1, get_consensus_label)
}


```






