---
title: "Vedo2Project scRNA b7 ratios"
author: "Paola Pibiri"
output: html_document
---

```{r header, include=FALSE}
############################################################################################################################
# Authors: Paola Pibiri
# Name: Vedo2Project_scRNA_b7_ratios.Rmd
# Function: check the cell-cell communications between 1 receiver cell and multiple sender cells 
############################################################################################################################
```



```{r libraries, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

####################
# libraries        #
####################

library(dplyr)
library(Seurat)

```



```{r function, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#######################################
# Functions                           #
#######################################

add_lower_classification <- function(seurat_object, reclassification_mapping, mapping_original_column, mapping_reclass_column, metadata_original_column, metadata_reclassification_column){
  # add the new column
  seurat_object@meta.data[[metadata_reclassification_column]] <- NA
  # get each cell type in the data
  metadata_original_cts <- unique(seurat_object@meta.data[[metadata_original_column]])
  # and the originals in the mapping
  reclassification_original_cts <- unique(reclassification_mapping[[mapping_original_column]])
  # we can only map what is present in both
  originals_both <- intersect(metadata_original_cts, reclassification_original_cts)
  # check what is missing
  only_metadata <- setdiff(metadata_original_cts, reclassification_original_cts)
  only_mapping <- setdiff(reclassification_original_cts, metadata_original_cts)
  # warn what is missing
  if(length(only_metadata) > 0){
    print('some celltypes only in metadata')
    print(only_metadata)
  }
  if(length(only_mapping) > 0){
    print('some celltypes only in remapping ')
    print(only_mapping)
  }
  # check each cell type
  for(celltype_original in originals_both){
    # get the appropriate remapping
    celltype_remapped <- reclassification_mapping[reclassification_mapping[[mapping_original_column]] == celltype_original, mapping_reclass_column]
    # now remap in the metadata
    seurat_object@meta.data[seurat_object@meta.data[[metadata_original_column]] == celltype_original, metadata_reclassification_column] <- celltype_remapped
  }
  return(seurat_object)
}
```


## Proportion of B7 pos cells in scRNAseq PBMCs data 
```{r pbmc scrnaseq, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

## time points used for the analysis 
before.treatment <- c('T0', 'T1')
after.treatment <- c('T3', 'T4')


#read the object for the low granularity in pbmc
pbmc <- readRDS('/groups/umcg-weersma/tmp01/projects/vedopredict2/pbmc/ongoing/cell_type_classification/objects/vedo2_pbmc_merged_clustered_azi.rds')

#read the object for the low granularity mapping
pbmc_mapping <- read.table('/groups/umcg-weersma/tmp01/projects/vedopredict2/pbmc/ongoing/cell_type_classification/vedo2_celltype_lowres_remapping_pbmc2.tsv', header= T, sep='\t')
# create a column with lower granularity
pbmc <- add_lower_classification(pbmc, pbmc_mapping, 'celltype_pbmc', 'lowres', 'azimuth_l2', 'cell_type_lower')  

#check their abundance separately --> b7
feature_imm_b7_pbmc <- FeaturePlot(pbmc, features=c('ITGB7'))

pbmc@meta.data[pbmc@assays$RNA@data['ITGB7', ] >= 0.5, 'b7_positive'] <- T
pbmc@meta.data[pbmc@assays$RNA@data['ITGB7', ] < 0.5, 'b7_positive'] <- F

b7_pos_pbmc <- DimPlot(pbmc, group.by = 'b7_positive') + scale_color_manual(values=c('gray', 'blue'))

# run for b7 the function for checking the positive cell ratio in biopsy 
pbmc_NR <- pbmc[ ,pbmc@meta.data$PGA_response == 'no' & pbmc@meta.data$timepoint %in% before.treatment]
pbmc_R <- pbmc[ ,pbmc@meta.data$PGA_response == 'yes' & pbmc@meta.data$timepoint %in% before.treatment]


#calculate the proportion of PBMC B7 positive cells
b7_positive_list_pbmc <- list()

pbmc_subset <- pbmc_R

for (patient in unique(pbmc_subset@meta.data$patient)) {
  for (celltype in unique(pbmc_subset@meta.data$cell_type_lower)) {
    total_patient_cell <- nrow(pbmc_subset@meta.data[pbmc_subset@meta.data$patient == patient & pbmc_subset@meta.data$cell_type_lower == celltype, ])
    total_patient_cell_positive <- nrow(pbmc_subset@meta.data[pbmc_subset@meta.data$patient == patient & pbmc_subset@meta.data$cell_type_lower == celltype & pbmc_subset@meta.data$b7_positive == T, ])
    proportion <- total_patient_cell_positive/total_patient_cell
    b7_positive_list_pbmc[[paste0(patient, "_", celltype)]] <- data.frame(patient = patient, celltype = celltype, proportion = proportion, tot = total_patient_cell, pos = total_patient_cell_positive)
  }
}

b7_positive_df_pbmc <- do.call('rbind', b7_positive_list_pbmc)

write.table(b7_positive_df_pbmc, file= '/groups/umcg-weersma/tmp01/projects/vedopredict2/pbmc/ongoing/cell_type_classification/b7_positive_df_pbmc_R_t0.tsv', row.names = F, col.names = T, sep = '\t')


```


## Proportion of B7 pos cells in scRNAseq Biopsy data 
```{r biopsy scrnaseq, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

## time points used for the analysis 
before.treatment <- 'T0'
after.treatment <- 'T4'

#load the total object
vedo2 <- readRDS('/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/seurat_preprocess_samples/objects/vedo2_remerged_resct_20230215.rds')
set.seed(0)
vedo2<- RunPCA(vedo2)
vedo2 <- RunUMAP(vedo2, dims = 1:30)


#read the object for the low granularity mapping
biopsy_mapping <- read.table('/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/cell_type_classification/vedo2_celltype_lowres_remapping_biopsy2.tsv', header = T, sep='\t')
# create a column with lower granularity
vedo2 <- add_lower_classification(vedo2, biopsy_mapping, 'celltype_final', 'lowres', 'celltype_final', 'cell_type_lower')

#create a subset of the vedo2 object containing only the immune compartment
vedo2_immune <- vedo2[ ,vedo2@meta.data$compartment == 'immune']

#check their abundance separately --> b7
feature_imm_b7 <- FeaturePlot(vedo2_immune, features=c('ITGB7'))

vedo2_immune@meta.data[vedo2_immune@assays$RNA@data['ITGB7', ] >= 0.5, 'b7_positive'] <- T
vedo2_immune@meta.data[vedo2_immune@assays$RNA@data['ITGB7', ] < 0.5, 'b7_positive'] <- F

b7_pos <- DimPlot(vedo2_immune, group.by = 'b7_positive') + scale_color_manual(values=c('gray', 'blue'))


# run for b7 the function for checking the positive cell ratio in biopsy 
vedo2_NR <- vedo2_immune[ ,vedo2_immune@meta.data$PGA._resp == 'no' & vedo2_immune@meta.data$timepoint == before.treatment]
vedo2_R <- vedo2_immune[ ,vedo2_immune@meta.data$PGA._resp == 'yes' & vedo2_immune@meta.data$timepoint == before.treatment]


#calculate the biopsy B7 proportion
b7_positive_list <- list()

vedo2_subset <- vedo2_R

for (patient in unique(vedo2_subset@meta.data$patient)) {
  for (celltype in unique(vedo2_subset@meta.data$cell_type_lower)) {
    total_patient_cell <- nrow(vedo2_subset@meta.data[vedo2_subset@meta.data$patient == patient & vedo2_subset@meta.data$cell_type_lower == celltype, ])
    total_patient_cell_positive <- nrow(vedo2_subset@meta.data[vedo2_subset@meta.data$patient == patient & vedo2_subset@meta.data$cell_type_lower == celltype & vedo2_subset@meta.data$b7_positive == T, ])
    proportion <- total_patient_cell_positive/total_patient_cell
    b7_positive_list[[paste0(patient, "_", celltype)]] <- data.frame(patient = patient, celltype = celltype, proportion = proportion, tot = total_patient_cell, pos = total_patient_cell_positive)
  }
}

b7_positive_df_biopsy <- do.call('rbind', b7_positive_list)

write.table(b7_positive_df_biopsy, file= '/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/cell_type_classification/b7_positive_df_biopsy_R_t0.tsv', row.names = F, col.names = T, sep = '\t')

```



```{r read input files, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#read the 4 tables for biopsy and the 4 tables for pbmc
biopsy_R_t0 <- read.table('/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/cell_type_classification/b7_positive_df_biopsy_R_t0.tsv', sep='\t',  header = T)
biopsy_R_t4 <- read.table('/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/cell_type_classification/b7_positive_df_biopsy_R_t4.tsv', sep='\t',  header = T)
biopsy_NR_t0 <- read.table('/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/cell_type_classification/b7_positive_df_biopsy_NR_t0.tsv', sep='\t',  header = T)
biopsy_NR_t4 <- read.table('/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/cell_type_classification/b7_positive_df_biopsy_NR_t4.tsv', sep='\t',  header = T)

pbmc_R_t0 <- read.table('/groups/umcg-weersma/tmp01/projects/vedopredict2/pbmc/ongoing/cell_type_classification/b7_positive_df_pbmc_R_t0.tsv', sep='\t',  header = T)
pbmc_R_t4 <- read.table('/groups/umcg-weersma/tmp01/projects/vedopredict2/pbmc/ongoing/cell_type_classification/b7_positive_df_pbmc_R_t4.tsv', sep='\t',  header = T)
pbmc_NR_t0 <- read.table('/groups/umcg-weersma/tmp01/projects/vedopredict2/pbmc/ongoing/cell_type_classification/b7_positive_df_pbmc_NR_t0.tsv', sep='\t',  header = T)
pbmc_NR_t4 <- read.table('/groups/umcg-weersma/tmp01/projects/vedopredict2/pbmc/ongoing/cell_type_classification/b7_positive_df_pbmc_NR_t4.tsv', sep='\t',  header = T)

```


## Combine the results of Responders at T0 and calculate -log10 ratios
```{r R T0, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

####in R at T0
#rename the columns 
colnames(biopsy_R_t0) <- c('patient', 'celltype', 'proportion_biopsy', 'tot_biopsy', 'pos_biopsy')
colnames(pbmc_R_t0) <- c('patient', 'celltype', 'proportion_pbmc', 'tot_pbmc', 'pos_pbmc')

#merge the two files
b7_merged <- merge(biopsy_R_t0,pbmc_R_t0,by= c("patient", "celltype"))

#save the files
write.table(b7_merged, file= '/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/R_t0_b7_pos_pbmc_vs_biopsy_all.tsv', sep='\t', row.names = F, col.names = T)

#filter for only relevant columns 
b7_merged_filtered <- b7_merged[ , c('patient', 'celltype', 'proportion_biopsy', 'proportion_pbmc')]


#calculate the ratio and the -log2 ratio
b7_merged_filtered$ratio <- b7_merged_filtered$proportion_pbmc/b7_merged_filtered$proportion_biopsy
b7_merged_filtered$ratio_log <- -log2(b7_merged_filtered$proportion_pbmc/b7_merged_filtered$proportion_biopsy)

#filter our the not complete cases
b7_merged_filtered <- b7_merged_filtered[!(b7_merged_filtered$ratio %in% c('NaN', 'Inf')), ]
b7_merged_filtered <- b7_merged_filtered[!(b7_merged_filtered$ratio_log %in% c('NaN', 'Inf')), ]
b7_merged_filtered <- b7_merged_filtered[complete.cases(b7_merged_filtered), ]

#save the final table
write.table(b7_merged_filtered, file= '/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/R_t0_ratio_b7_pos_pbmc_vs_biopsy.tsv', sep='\t', row.names = F, col.names = T)


```


## Combine the results of Non-Responders at T0 and calculate -log10 ratios
```{r NR T0, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#rename the columns
colnames(biopsy_NR_t0) <- c('patient', 'celltype', 'proportion_biopsy', 'tot_biopsy', 'pos_biopsy')
colnames(pbmc_NR_t0) <- c('patient', 'celltype', 'proportion_pbmc', 'tot_pbmc', 'pos_pbmc')

#merge the two files
b7_merged <- merge(biopsy_NR_t0,pbmc_NR_t0,by= c("patient", "celltype"))

#save it
write.table(b7_merged, file= '/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/NR_t0_b7_pos_pbmc_vs_biopsy_all.tsv', sep='\t', row.names = F, col.names = T)

#filter for only relevant columns
b7_merged_filtered <- b7_merged[ , c('patient', 'celltype', 'proportion_biopsy', 'proportion_pbmc')]

#calculate ratio and -loag2 ratio
b7_merged_filtered$ratio <- b7_merged_filtered$proportion_pbmc/b7_merged_filtered$proportion_biopsy
b7_merged_filtered$ratio_log <- -log2(b7_merged_filtered$proportion_pbmc/b7_merged_filtered$proportion_biopsy)

#filter out not complete cases
b7_merged_filtered <- b7_merged_filtered[!(b7_merged_filtered$ratio %in% c('NaN', 'Inf')), ]
b7_merged_filtered <- b7_merged_filtered[!(b7_merged_filtered$ratio_log %in% c('NaN', 'Inf')), ]
b7_merged_filtered <- b7_merged_filtered[b7_merged_filtered(b7_merged_filtered), ]

#save table
write.table(b7_merged_filtered, file= '/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/NR_t0_ratio_b7_pos_pbmc_vs_biopsy.tsv', sep='\t', row.names = F, col.names = T)
```


## Combine the results of Responders at T4 and calculate -log10 ratios
```{r R T4, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#rename 
colnames(biopsy_R_t4) <- c('patient', 'celltype', 'proportion_biopsy', 'tot_biopsy', 'pos_biopsy')
colnames(pbmc_R_t4) <- c('patient', 'celltype', 'proportion_pbmc', 'tot_pbmc', 'pos_pbmc')

#merge the two files
b7_merged <- merge(biopsy_R_t4,pbmc_R_t4,by= c("patient", "celltype"))

#save the table
write.table(b7_merged, file= '/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/R_t4_b7_pos_pbmc_vs_biopsy_all.tsv', sep='\t', row.names = F, col.names = T)

#filter for only relevant columns
b7_merged_filtered <- b7_merged[ , c('patient', 'celltype', 'proportion_biopsy', 'proportion_pbmc')]



#calculate the ratio and the -log2 ratio
b7_merged_filtered$ratio <- b7_merged_filtered$proportion_pbmc/b7_merged_filtered$proportion_biopsy
b7_merged_filtered$ratio_log <- -log2(b7_merged_filtered$proportion_pbmc/b7_merged_filtered$proportion_biopsy)

#filter out not complete cases 
b7_merged_filtered <- b7_merged_filtered[!(b7_merged_filtered$ratio %in% c('NaN', 'Inf')), ]
b7_merged_filtered <- b7_merged_filtered[!(b7_merged_filtered$ratio_log %in% c('NaN', 'Inf')), ]
b7_merged_filtered <- b7_merged_filtered[complete.cases(b7_merged_filtered), ]

#save the table
write.table(b7_merged_filtered, file= '/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/R_t4_ratio_b7_pos_pbmc_vs_biopsy.tsv', sep='\t', row.names = F, col.names = T)

```



## Combine the results of Non-Responders at T4 and calculate -log10 ratios
```{r NR T4, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#rename the columns
colnames(biopsy_NR_t4) <- c('patient', 'celltype', 'proportion_biopsy', 'tot_biopsy', 'pos_biopsy')
colnames(pbmc_NR_t4) <- c('patient', 'celltype', 'proportion_pbmc', 'tot_pbmc', 'pos_pbmc')

#merge the two files
b7_merged <- merge(biopsy_NR_t4,pbmc_NR_t4,by= c("patient", "celltype"))

#save the table
write.table(b7_merged, file= '/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/NR_t4_b7_pos_pbmc_vs_biopsy_all.tsv', sep='\t', row.names = F, col.names = T)

#filter for only relevant columns
b7_merged_filtered <- b7_merged[ , c('patient', 'celltype', 'proportion_biopsy', 'proportion_pbmc')]


#calculate ratio and -log2 ratio
b7_merged_filtered$ratio <- b7_merged_filtered$proportion_pbmc/b7_merged_filtered$proportion_biopsy
b7_merged_filtered$ratio_log <- -log2(b7_merged_filtered$proportion_pbmc/b7_merged_filtered$proportion_biopsy)

#filter out not complete cases 
b7_merged_filtered <- b7_merged_filtered[!(b7_merged_filtered$ratio %in% c('NaN', 'Inf')), ]
b7_merged_filtered <- b7_merged_filtered[!(b7_merged_filtered$ratio_log %in% c('NaN', 'Inf')), ]
b7_merged_filtered <- b7_merged_filtered[complete.cases(b7_merged_filtered), ]


write.table(b7_merged_filtered, file= '/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/NR_t4_ratio_b7_pos_pbmc_vs_biopsy.tsv', sep='\t', row.names = F, col.names = T)
```


## Merging together the 4 files with the 4 conditions
```{r merging, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#read the final files in which pbmc and biopsy are merged 

R_t0 <- read.table('/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/R_t0_ratio_b7_pos_pbmc_vs_biopsy.tsv', sep='\t',  header = T)
R_t4 <- read.table('/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/R_t4_ratio_b7_pos_pbmc_vs_biopsy.tsv', sep='\t',  header = T)
NR_t0 <- read.table('/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/NR_t0_ratio_b7_pos_pbmc_vs_biopsy.tsv', sep='\t',  header = T)
NR_t4 <- read.table('/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/NR_t4_ratio_b7_pos_pbmc_vs_biopsy.tsv', sep='\t',  header = T)

#create a condition column
R_t0$condition <- "R-T0"
R_t4$condition <- "R-T4"
NR_t0$condition <- "NR-T0"
NR_t4$condition <- "NR-T4"

#merge dataframes
df <- do.call("rbind", list(R_t0, R_t4, NR_t0, NR_t4))
write.table(df, file= '/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/merged_df_proportion_b7.tsv', sep='\t', row.names = F, col.names = T)
```


## Wilcoxon Rank Sum Test for all the conditions
```{r merging, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#filter the datasets
R_t0_filtered <- R_t0[ , c('patient', 'celltype', 'ratio_log')]
R_t4_filtered <- R_t4[ , c('patient', 'celltype', 'ratio_log')]
NR_t0_filtered <- NR_t0[ , c('patient', 'celltype', 'ratio_log')]
NR_t4_filtered <- NR_t4[ , c('patient', 'celltype', 'ratio_log')]



# calculate wilcoxon in NR t0 vs R t0
wilcoxon_test <- NULL
for(celltype in unique(c(NR_t0_filtered[['celltype']], R_t0_filtered[['celltype']]))){
  x <- NR_t0_filtered[NR_t0_filtered['celltype'] == celltype, 'ratio_log']
  y <- R_t0_filtered[R_t0_filtered['celltype'] == celltype, 'ratio_log']
  if (length(x) > 0 & length(y) >0) {
    result_wilcox <- wilcox.test(x, y, paired = FALSE, exact = FALSE)
    P_value <- result_wilcox[["p.value"]]
    wilcoxon_row <- data.frame(celltype=c(celltype), p=c(P_value))
    if(is.null(wilcoxon_test)){
      wilcoxon_test <- wilcoxon_row
    }
    else{
      wilcoxon_test <- rbind(wilcoxon_test, wilcoxon_row)
    }
  } else {
    print(paste0('skipping,', celltype , ' due to not enough observations: NR ', length(x), ' R ', length(y)))
  }
}

write.table(wilcoxon_test, file= '/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/wilcoxon_results_NR_vs_R_t0_b7_log_ratio.tsv', sep='\t', col.names = T, row.names = F)



# calculate wilcoxon in NR t4 vs R t4
wilcoxon_test <- NULL
for(celltype in unique(c(NR_t4_filtered[['celltype']], R_t4_filtered[['celltype']]))){
  x <- NR_t4_filtered[NR_t4_filtered['celltype'] == celltype, 'ratio_log']
  y <- R_t4_filtered[R_t4_filtered['celltype'] == celltype, 'ratio_log']
  if (length(x) > 0 & length(y) >0) {
    result_wilcox <- wilcox.test(x, y, paired = FALSE, exact = FALSE)
    P_value <- result_wilcox[["p.value"]]
    wilcoxon_row <- data.frame(celltype=c(celltype), p=c(P_value))
    if(is.null(wilcoxon_test)){
      wilcoxon_test <- wilcoxon_row
    }
    else{
      wilcoxon_test <- rbind(wilcoxon_test, wilcoxon_row)
    }
  } else {
    print(paste0('skipping,', celltype , ' due to not enough observations: NR ', length(x), ' R ', length(y)))
  }
}


write.table(wilcoxon_test, file= '/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/wilcoxon_results_NR_vs_R_t4_b7_log_ratio.tsv', sep='\t', col.names = T, row.names = F)


# Check R t4 vs t0
wilcoxon_test <- NULL
for(celltype in unique(c(R_t4_filtered[['celltype']], R_t0_filtered[['celltype']]))){
  x <- R_t4_filtered[R_t4_filtered['celltype'] == celltype, 'ratio_log']
  y <- R_t0_filtered[R_t0_filtered['celltype'] == celltype, 'ratio_log']
  if (length(x) > 0 & length(y) >0) {
    result_wilcox <- wilcox.test(x, y, paired = FALSE, exact = FALSE)
    P_value <- result_wilcox[["p.value"]]
    wilcoxon_row <- data.frame(celltype=c(celltype), p=c(P_value))
    if(is.null(wilcoxon_test)){
      wilcoxon_test <- wilcoxon_row
    }
    else{
      wilcoxon_test <- rbind(wilcoxon_test, wilcoxon_row)
    }
  } else {
    print(paste0('skipping,', celltype , ' due to not enough observations: NR ', length(x), ' R ', length(y)))
  }
}

write.table(wilcoxon_test, file= '/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/wilcoxon_results_t4_vs_t0_R_b7_log_ratio.tsv', sep='\t', col.names = T, row.names = F)



# Check NR t4 vs t0
wilcoxon_test <- NULL
for(celltype in unique(c(NR_t4_filtered[['celltype']], NR_t0_filtered[['celltype']]))){
  x <- NR_t4_filtered[NR_t4_filtered['celltype'] == celltype, 'ratio_log']
  y <- NR_t0_filtered[NR_t0_filtered['celltype'] == celltype, 'ratio_log']
  if (length(x) > 0 & length(y) >0) {
    result_wilcox <- wilcox.test(x, y, paired = FALSE, exact = FALSE)
    P_value <- result_wilcox[["p.value"]]
    wilcoxon_row <- data.frame(celltype=c(celltype), p=c(P_value))
    if(is.null(wilcoxon_test)){
      wilcoxon_test <- wilcoxon_row
    }
    else{
      wilcoxon_test <- rbind(wilcoxon_test, wilcoxon_row)
    }
  } else {
    print(paste0('skipping,', celltype , ' due to not enough observations: NR ', length(x), ' R ', length(y)))
  }
}

write.table(wilcoxon_test, file= '/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/wilcoxon_results_t4_vs_t0_NR_b7_log_ratio.tsv', sep='\t', col.names = T, row.names = F)

# read all the wilcoxon files and merged them
NR_vs_R_t0 <- read.table('/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/wilcoxon_results_NR_vs_R_t0_b7_log_ratio.tsv', sep='\t',  header = T)
NR_vs_R_t4 <- read.table('/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/wilcoxon_results_NR_vs_R_t4_b7_log_ratio.tsv', sep='\t',  header = T)
t4_vs_t0_R <- read.table('/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/wilcoxon_results_t4_vs_t0_R_b7_log_ratio.tsv', sep='\t',  header = T)
t4_vs_t0_NR <- read.table('/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/wilcoxon_results_t4_vs_t0_NR_b7_log_ratio.tsv', sep='\t',  header = T)

#add the column of the condition
NR_vs_R_t0$comparison <- "NR-T0 vs R-T0"
NR_vs_R_t4$comparison <- "NR-T4 vs R-T4"
t4_vs_t0_R$comparison <- "R-T4 vs R-T0"
t4_vs_t0_NR$comparison <- "NR-T4 vs NR-T0"

#merge dataftames
df_wilcoxon <- do.call("rbind", list(NR_vs_R_t0, NR_vs_R_t4, t4_vs_t0_R, t4_vs_t0_NR))

write.table(df_wilcoxon, file= '/groups/umcg-weersma/tmp01/projects/vedopredict2/biopsies/ongoing/new_results_b7_proportion/wilcoxon_results_final_table.tsv', sep='\t', col.names = T, row.names = F)

```


## Boxplots
```{r boxplot, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# melt the 4 datasets 
R_t0_melt <-  melt(R_t0_filtered)
R_t4_melt <-  melt(R_t4_filtered)
NR_t0_melt <-  melt(NR_t0_filtered)
NR_t4_melt <-  melt(NR_t4_filtered)

# add the condition column for each dataset
R_t0_melt$condition <- "R-T0"
R_t4_melt$condition <- "R-T4"
NR_t0_melt$condition <- "NR-T0"
NR_t4_melt$condition <- "NR-T4"

# run the rbind function across datasets 
merged <- do.call("rbind", list(R_t0_melt, R_t4_melt, NR_t0_melt, NR_t4_melt))


#remove NA, Inf or -inf 

merged_filtered <- merged

merged_filtered <- merged_filtered[!is.na(merged_filtered$value), ]
merged_filtered <- merged_filtered[is.finite(merged_filtered$value), ]

#create the plot
merged_filtered$condition <- factor(merged_filtered$condition, levels = c("R-T0", "NR-T0", "R-T4", "NR-T4"))

ggplot(merged_filtered, aes(x=condition, y=value, fill=condition)) + 
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(size = 0.8, width = 0.15, alpha = 0.5) +
  facet_wrap( ~ celltype, scales="free") +
  theme(panel.background = element_rect(fill = 'white', color='black')) +
  theme(strip.background = element_rect(fill="white"),
        strip.text = element_text(colour="black"), panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                                                                   colour = "white")) +
  
  scale_fill_manual(values = list('R-T0' = 'white', 'R-T4' = 'white', 'NR-T0' = 'gray', 'NR-T4' = 'gray'))
```








